Equilibria and Convergence of Fictitious Play on Network Aggregative Games

### 2021-05-21 Documents/FP in NAGs/neurips_2021.tex
Updated 2021 doc

Multi Agent Learning \cite{Schwartz} requires a number of agents to adapt in an environment where
each agent must adapt in response to the behaviour of the other agents. This leads to a
fundamentally non-stationary problem, which presents a challenge to designing effective learning
policies. Even when there are a small number of agents in the game, learning has been shown to lead
to non-stationary, and even chaotic behaviour \cite{SatoChaos}, and this problem becomes even more
pronounced as the number of agents increase \cite{Sanders}. In this light, it would seem that, when there is a large population of agents, it is almost impossible to understand the long term behaviour of multi agent learning \cite{PiliourasChaoticMaps}. This, of course, poses a problem for AI safety, for which it is necessary that learning systems are designed so that their ultimate behaviour is guaranteed to satisfy some predetermined goals.

To deal with this problem, one must attempt to reduce the many player game to something which is
tractable. A number of reductions have been proposed, most notably \emph{Mean Field} Games
and \emph{Aggregative} Games. The former makes the assumption of an infinite
number of agents, so that the population can be represented through a distribution over players'
states \cite{CainesPaper} . Every agent then updates their action profiles depending on this
distribution. In the latter each agent plays considers a real valued function which is a convex
combination of the states of the other agents. In both formulations, a many player game is reduced
to the set of two player games which each agent is involved in.

Both formulations have been the object of rigorous study and it has been shown that agents reach an equilibrium when learning on such games (c.f. \cite{MFGLearningPapers} for Mean Field Games and
\cite{Aggregative Papers} for Aggregative Games). However, both also present a fundamental limitation.
Namely, they both require that agents have access to the action profiles of the entire population.
This could be through communication with all other agents, or through the intervention of a central
coordinator who is able to the entire population. Whilst recent work aims to relax this assumption
through the introduction of noise \cite{MFG-FP} or partial observability \cite{AAMASPaper}, the
requirement that each agent updates their actions based on the entire population is rather strong.

@@

  We first show that a Nash Equilibrium exists and that FP admits
  solutions in this setting. In particular, we study zero-sum games and show that
  FP converges to a fixed point, which for a network without any
  self-loops corresponds to a Nash Equilibrium. In addition, we find that, for games
  which are not zero-sum, agents following FP are able to
  achieve no regret.

@@

The NA game is a recent introduction \cite{Parise2015} which presents a variant of aggregative games
by adding an underlying structure to the population. Since its introduction, distributed algorithms
have been built with the aim of finding NE. In particular, \cite{Parise2015, Parise2020} consider
the case in which payoffs are given by Lipschitz functions with unique minimisers and apply standard
topological fixed point towards designing algorithms which will converge to the NE. We note that the
\emph{Mann Iteration}, one of the algorithms applied in \cite{Parise2020} shares a remarkably
similar form with the discrete variant of FP. To the best of our knowledge, this link has not been
explored in the literature and is an avenue for research which we are currently pursuing. Another
common algorithm for distributed NE seeking is the projected gradient (resp. subgradient) dynamics
which is explored in \cite{Zhang2020} (resp. \cite{Shokri2020, Shokri2021}). In all of these works,
the cost function is assumed to be convex, and therefore have a unique minimiser. In fact this is a
common assumption in works concerning NA Games \cite{Zhu2021, Lei2020}, which we believe is due to
its ubiquity in control settings. We have not yet come across works which consider NA games from the
point of view of payoff matrices, which are more common in multi agent learning settings.
Furthermore, to the best of our knowledge, this is the first work which introduces the application
of a strict learning algorithm in the NA setting.

@@

Fictitious Play was introduced by Brown in 1951 \cite{BrownPublished, BrownUnpublished} as a
'natural' way in which to approximate the Nash Equilibrium in zero sum games. It was known at the
time that a discrete variant of the algorithm would converge to the NE in two player zero-sum games
\cite{Robinson}. Since then a number of proofs of convergence were determined in 2 player games
\cite{Miyasawa, Polak, Berger, Monderer and Sela, Monderer and Shapley}, as well as corresponding
convergence rates \cite{Harris, Shapiro}. Many results hold for both the discrete and continuous
variants of FP. Their relation was finally solved by Hofbauer \cite{Hofbauer}, who showed that
attractors of the discrete time process are also attractors for the continuous time variant. As
such, most of the analysis on FP focuses on the continuous time variant \cite{Ostrovski}. It was
later shown in \cite{Shapley} that there is a family of games (now called the \emph{Shapley family})
for which FP does not converge at all but rather shows periodic behaviour. This led to research
which looks at this non convergent behaviour, specifically showing the existence of more periodic
orbits as well as chaotic behaviour \cite{Sparrow}. However, the work looking at fictitious play
where there are more than two agents is sparse. \cite{Sela} considers a game in which a multi-player
game is decomposed by requiring that each agent engage in a two person game against each of the
opponent. Their payoff is given by the sum of payoffs in all of these subgames. It was found that,
if this game is zero sum, then FP will converge. Similar results for more than two players were
found for games with identical interests in \cite{MondererShapley1994}. In \cite{FPMFG}, the action
of FP was considered in a Mean Field Game, and it was found again that the algorithm converges for
zero sum games. The most general result, and the one most similar to our own, was found by Ewerhart
and Valkanova in 2020 \cite{Ewerhart} in which they showed that fictitious play converges in network
games, where each agent is engaged in a two-player game with each of their neighbours. The rate of
convergence was found to be independent of the network size.  Our work extends the analysis of FP in
multiplayer games by considering its action in NA Games, so that the agents do not play individual
games against each of their neighbours but rather a single game against the aggregate of their
neighbours. In particular we also extend a result for two player games found by Ostrovski and van
Strein \cite{Payoff Performance} which showed that FP achieves no-regret to the multi-player
setting.

\section{Preliminaries}
\subsection{Network Aggregative Games}
\label{sec::NAG}

@@

  In Figure \ref{fig::3PlayerChainNoise} we revisit the Three Player Chain of Section
  \ref{sec::NonConv}, now under the influence of additive noise. For the sake of brevity, we only
  display the distance to the Nash Equilibrium of the first player's action, since the other agents
  behave in the same way. It can be seen that a small amount of noise has the effect of decreasing
  the size of the periodic orbit. However, as $\gamma$
  is increased to $0.5$, the algorithm seems to exhibit convergence to the NE. The implication is
  that the addition of noise may cause periodic behaviour to break and lead to the Nash Equilibrium.
  An interesting point to note is that this behaviour is in stark contrast to the replicator
  dynamic (RD) \cite{Maynard-Smith}, another adaptive algorithm linked to multi agent learning
  \cite{CyclesAdversarialLearning}. In \cite{Imhof} and \cite{Galla}, it was found that the
  introduction of random mutations can remove convergent behaviour and instead lead to periodicity.

@@

Finally, our work has a strong impact on the study of the Network Aggregative Game, which has strong
applications in multi agent control \cite{GrammaticoAggregativeControl}. We 
believe that our work makes a strong step towards ensuring that systems which learn and adapt on
NA games maintain stability and, therefore, can be considered safe.

>>>>

Multi-Agent Learning \cite{Schwartz} requires a number of agents to adapt in an environment where
each agent must adapt in response to the behaviour of the other agents. This leads to a
fundamentally non-stationary problem, which presents a challenge to designing effective learning
policies. Even when there are a small number of agents in the game, learning has been shown to lead
to non-stationary, and even chaotic behaviour \cite{SatoChaos}, and this problem becomes even more
pronounced as the number of agents increase \cite{Sanders}. In this light, it would seem that, when there is a large population of agents, it is almost impossible to understand the long term behaviour of multi-agent learning \cite{PiliourasChaoticMaps}. This, of course, poses a problem for AI safety, for which it is desirable that learning systems are designed so that their ultimate behaviour is guaranteed to satisfy some predetermined goals.

To deal with this problem, one must attempt to reduce the many player game to something which is
tractable. A number of reductions have been proposed, most notably \emph{Mean Field} Games
and \emph{Aggregative} Games. The former makes the assumption of an infinite
number of agents, so that the population can be represented through a distribution over players'
states \cite{CainesPaper} . Every agent then updates their action profiles depending on this
distribution. In the latter each agent considers a real valued function which is a convex
combination of the states of the other agents. Both formulation allows for a many player game to be reduced to a set of two-player games.

Both formulations have been the object of rigorous study and it has been shown that agents reach an equilibrium when learning on such games (c.f. \cite{MFGLearningPapers} for Mean Field Games and
\cite{Aggregative Papers} for Aggregative Games). However, both also present a fundamental limitation.
Namely, they both require that agents have access to the action profiles of the entire population.
This could be through communication with all other agents, or through the intervention of a central
coordinator who is able view the entire population. Whilst recent work aims to relax this assumption
through the introduction of noise \cite{MFG-FP} or partial observability \cite{AAMASPaper}, the
requirement that each agent updates their actions based on the entire population is rather strong and not always supported by empirical evidence.

@@

  We first show that, in aggregative games played with FP, a Nash Equilibrium exists and that FP admits
  solutions in this setting. In particular, we study zero-sum games and show that
  FP converges to a fixed point, which for a network without any
  self-loops corresponds to a Nash Equilibrium. In addition, we find that, for games
  which are not zero-sum, agents following FP are able to
  achieve no regret.

@@

The NA game is a recent introduction \cite{Parise2015} which presents a variant of aggregative games
by adding an underlying structure to the population. Since its introduction, distributed algorithms
have been built with the aim of finding NE. In particular, \cite{Parise2015, Parise2020} consider
the case in which payoffs are given by Lipschitz functions with unique minimisers and apply standard
topological fixed point towards designing algorithms which will converge to the NE. We note that the
\emph{Mann Iteration}, one of the algorithms applied in \cite{Parise2020} shares a remarkably
similar form with the discrete variant of FP. To the best of our knowledge, this link has not been
explored in the literature and is an avenue for research which we are currently pursuing. Another
common algorithm for distributed NE seeking is the projected gradient (resp. subgradient) dynamics
which is explored in \cite{Zhang2020} (resp. \cite{Shokri2020, Shokri2021}). In all of these works,
the cost function is assumed to be convex, and therefore have a unique minimiser. In fact this is a
common assumption in works concerning NA Games \cite{Zhu2021, Lei2020}, which we believe is due to
its ubiquity in control settings. We have not yet come across works which consider NA games from the
point of view of payoff matrices, which are more common in multi-agent learning settings.
Furthermore, to the best of our knowledge, this is the first work which introduces the application
of a learning algorithm in the NA setting.

@@

Fictitious Play was introduced by Brown in 1951 \cite{BrownPublished, BrownUnpublished} as a
'natural' way in which to approximate the Nash Equilibrium in zero sum games. It was known at the
time that a discrete variant of the algorithm would converge to the NE in two player zero-sum games
\cite{Robinson}. Since then a number of proofs of convergence were determined in 2 player games
\cite{Miyasawa, Polak, Berger, Monderer and Sela, Monderer and Shapley}, as well as corresponding
convergence rates \cite{Harris, Shapiro}. Many results hold for both the discrete and continuous
variants of FP. Their relation was finally solved by Hofbauer \cite{Hofbauer}, who showed that
attractors of the discrete time process are also attractors for the continuous time variant. As
such, most of the analysis on FP focuses on the continuous time variant \cite{Ostrovski}. It was
later shown in \cite{Shapley} that there is a family of games (now called the \emph{Shapley family})
for which FP does not converge at all but rather shows periodic behaviour. However, the work looking at fictitious play
where there are more than two agents is sparse. \cite{Sela} considers a game in which a multi-player
game is decomposed by requiring that each agent engage in a two person game against each of the
opponent. Their payoff is given by the sum of payoffs in all of these subgames. It was found that,
if this game is zero sum, then FP will converge. Similar results for more than two players were
found for games where all agents share the same payoff in \cite{MondererShapley1994}. In \cite{FPMFG}, the action
of FP was considered in a Mean Field Game, and it was found again that the algorithm converges for
zero sum games. The most general result, and the one most similar to our own, was found by Ewerhart
and Valkanova in 2020 \cite{Ewerhart} in which they showed that fictitious play converges in network
games, where each agent is engaged in a two-player game with each of their neighbours. The rate of
convergence was found to be independent of the network size.  Our work extends the analysis of FP in
multiplayer games by considering its action in NA Games, so that the agents do not play individual
games against each of their neighbours but rather a single game against the aggregate of their
neighbours. In particular we also extend a result for two player games found by Ostrovski and van
Strein \cite{Payoff Performance} which showed that FP achieves no-regret to the multi-player
setting.

\section{Preliminaries}
In this section we introduce the network aggregative game framework, as well as defining FP on such a game.

\subsection{Network Aggregative Games}
\label{sec::NAG}

@@

  In Figure \ref{fig::3PlayerChainNoise} we revisit the Three Player Chain of Section
  \ref{sec::NonConv}, now under the influence of additive noise. For the sake of brevity, we only
  display the distance to the Nash Equilibrium of the first player's action, since the other agents
  behave in the same way. It can be seen that a small amount of noise has the effect of decreasing
  the size of the periodic orbit. However, as $\gamma$
  is increased to $0.5$, the algorithm seems to exhibit convergence to the NE. The implication is
  that the addition of noise may cause periodic behaviour to break and lead to the Nash Equilibrium.
  An interesting point to note is that this behaviour is in stark contrast to the replicator
  dynamic (RD) \cite{Maynard-Smith}, another adaptive algorithm linked to multi-agent learning
  \cite{CyclesAdversarialLearning}. In \cite{Imhof} and \cite{Galla}, it was found that the
  introduction of random mutations can remove convergent behaviour and instead lead to periodicity.

@@

Finally, our work has a strong impact on the study of the Network Aggregative Game, which has strong
applications in multi-agent control \cite{GrammaticoAggregativeControl}. We 
believe that our work makes a strong step towards ensuring that systems which learn and adapt on
NA games maintain stability and, therefore, can be considered safe.
